A geometric approach to linear and quadratic Chabauty

### 2020-02-27 MS.tex
Overview added and moved stuff

\usepackage{graphicx}
%\pagestyle{empty}
\usepackage{dsfont}
\usepackage{textcomp}
\usepackage{amsmath}
\usepackage{courier}
\usepackage[utf8]{inputenc}
\usepackage{graphicx}
\graphicspath{{images/}}
\usepackage{amsthm}
\usepackage{eucal}
\usepackage{amssymb}
%\usepackage{mdsymbol}
\usepackage{mathrsfs}
\usepackage{dsfont}
\usepackage[usenames,dvipsnames,svgnames,table]{xcolor}
\usepackage{caption}
\usepackage{tikz}
\usepackage{tikz-cd}
\usepackage{wrapfig,lipsum,booktabs}
\usepackage{caption}
\usepackage[square, longnamesfirst,numbers]{natbib}
\usepackage[ruled,linesnumbered]{algorithm2e}
\usepackage{todonotes}

@@

{\LARGE\bf 
A geometric approach to linear and quadratic Chabauty
}

@@

This Coleman-Chabauty method has been greatly generalised by Kim to so called non-abelian Chabauty. He interprets working in the Jacobian as dealing with the abelianised fundamental group of the curve $C$, and works with larger, non-abelian quotients of the fundamental group. Using quadratic Chabauty, a special case of non-abelian Chabauty, Balakrishnan, Dogra, M\"uller, Tuitman and Vonk famously calculated in May of 2019 all rational points of the ``cursed curve'', the modular curve $X_s(13)$. They use an endomorphism of the Jacobian, and do $p$-adic analysis on $p$-adic local height to accomplish this fact. \todo{Bas, is dit de goede samenvatting? Voornamelijk mijn vraag: is het nog steeds een specifiek geval van Coleman-Chabauty?}

Very recently, Edixhoven and Lido made available a preprint of their article ``Geometric quadratic Chabauty''. Their goal in this article is to make a more geometric approach to quadratic Chabauty, by working in a pullback $T$ of the Poincar\'e torsor of the Jacobian. They also abandon the Coleman method for Chabauty; instead they parametrise the map $T(\Z) \to T(\Z_p)$ with power series, and pull back equations for the curve along this map.

@@

\begin{defn}
Let $S$ be a scheme, $T \to U$ a morphism of schemes and $x : T \to S$ a $T$-point. We define $S(U)_x$ as the morphisms from $U$ to $S$ that, after precomposing with $T \to U$, give $x$.
\end{defn}

We want to find an upper bound for the cardinality of $C(\Z)_P$.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\newpage
\section{Linear Chabauty}
\label{section:linchab}

% Say something about smooth schemes over Z_p, and their Z_p points
\subsection{Points on a smooth scheme over \texorpdfstring{$\Z_p$}{Z\_p}}
\label{subs:smoothzppoints}
Let $X/\Z_p$ be a smooth scheme of relative dimension $d$, and let $x \in X(\F_p)$ be a point. Then $X(\Z_p)_x$ is in bijection with $\Z_p^d$. This bijection is given by choosing parameters at $x$; evaluating at $X(\Z_p)_x$ gives a bijection with $(p\Z_p)^d$, and then we divide by $p$. For putting up a nice framework to work in, we start with blowing up $X$ at $x$.
%TODO: give proof and define $\Z_p\angle{z_1,...,z_r}$ as the $p$-adic completion of $\Z_p[z_1,...,z_r]$, i.e. powerseries which are polynomials modulo every power of $p$.

@@

\subsection{From \texorpdfstring{$J(\Z)$}{J(Z)} to \texorpdfstring{$J(\Z_p)$}{J(Z\_p)}}
For $p > 2$, we know that the torsion of $J(\Z)$ injects into $J(\F_p)$, by Proposition 2.3 of \cite{pierre2000}. Hence for $0 \in J(\F_p)$, we know $J(\Z)_0$ is as a group isomorphic to $\Z^r$ with $r$ the Mordell-Weil rank. By assumption, we also know $J(\Z)_t$ is in bijection with $J(\Z)_0$, with the bijection giving by translating with a lift of $t$. By Subsection \ref{subs:smoothzppoints} we know $J(\Z_p)_t$ is in bijection with $\Z_p^{g}$, with the bijection given by evaluating parameters and dividing by $p$. Let $\kappa: \Z^r \to \Z_p^g$ be the map resulting from the inclusion $J(\Z)_t \to J(\Z_p)_t$. Then $\kappa$ turns out to have a special property.

@@

\subsubsection{From group schemes to formal groups}
We first recall the definition of a formal group. For this entire section, $R$ is a ring, and $x,y,z$ are vectors of variables.
\begin{defn}
\label{defn:formalgroup}
Let $n$ be a non-negative integer. Let $x,y,z$ be vectors of $n$ variables. An $n$-dimensional formal group is an element $F = (F_1,\dots,F_n) \in R[[x,y]]_0^n$ satisfying $F \equiv x + y \mod (x,y)^2$ and $F(F(x,y),z) = F(x,F(y,z))$. If furthermore $F(x,y) = F(y,x)$, this formal group is said to be commutative.
\end{defn}
\begin{exmp}
\label{exmp:gmformalgroup}
Take $n=1$ and $F(x,y) = x + y + xy = (1+x)(1+y)-1$, also known as the multiplicative formal group. This satisfies associativity as $F(F(x,y),z) = (1+x)(1+y)(1+z)-1 = F(x,F(y,z))$.
\end{exmp}
\begin{exmp}
\label{exmp:ganformalgroup}
For any $n$, we can take $F(x,y) = x+ y$, also known as the $n$-dimensional additive formal group.
\end{exmp}
[TODO: add text explaining the following lemma]
Note there is no mention of an inverse, but the following lemma, a formal version of the implicit function theorem, tells us that the inverse follows automatically from the definitions.
\begin{lem}
\label{lem:implicitfunction}
Let $x,y$ have length $n$. Let $F \in R[[x,y]]_0^n$ such that $F \equiv Ax + By \bmod (x,y)^2$ with $B \in \GL_n(R)$. Then there's a unique $\iota \in R[[x]]_0^n$ such that $F(x,\iota(x)) = 0$. 
\end{lem}
\begin{proof}
If $F$ is a polynomial, this is a special case of a multivariate version of Hensel's lemma, as in Corrolaire~2 of \cite[III,4.5]{BourbakiCA}, over $R[[x]]$ with maximal ideal $\m = (x)$, as the derivative matrix of $F(x,\iota)$ with respect to $\iota$ is $B$, which is invertible, and $\iota = 0$ gives a solution modulo $\m$. Now for $F$ any power series, let $F_j$ be the polynomial consisting of all terms in $F$ of degree at most $j$, and let $\iota_j$ be the unique power series in $R[[x]]_0^n$such that $F_j(x,\iota_j) = 0$. Note that both $\iota_j$ and $\iota_{j+1}$ are solutions to $F_j(x,\iota) \equiv 0 \mod \m^j$, so by the uniquess guaranteed by Hensel's lemma used over $R[[x]]/\m^j$, these must be equal. Hence they converge to $\iota \in R[[x]]_0^n$, which is the unique solution of $F(x,\iota(x)) = 0$.
\end{proof}
This has the following corollary about the inverse of a power series.
\begin{cor}
\label{lem:formalinverse}
Let $x$ have length $n$. Let $a \in R[[x]]_0^n$ with $a \equiv Px \mod x$ for some matrix $A\in \GL_n(R)$. Then there is a unique $b \in R[[x]]_0^n$ such that $a \circ b = b \circ a = x$.
\end{cor}
\begin{proof}
Let $F(x,y) = x - a(x)$. This satisfies the constraints of Lemma~\ref{lem:implicitfunction}, so we find a unique $b$ such that $a \circ b = x$. Applying Lemma~\ref{lem:implicitfunction} again gives a unique $c$ such that $b \circ c = x$. But then $a = a \circ (b \circ c) = (a \circ b) \circ c = c$ shows $a = c$ and hence we are done.
\end{proof}
So a formal group $F$ does by Lemma~\ref{lem:implicitfunction} indeed have a right inverse $\iota_F$. Also, by $F(0,0) = 0$ we have that $F(x,F(\iota_f,0)) = 0$ so $F(\iota_f,0)$ is in fact equal to $\iota_F$; as $\iota_F \equiv -x \mod (x)^2$, it has a formal inverse and hence $F(x,0) = x$, so $0$ is indeed a right unit. A standard argument now shows that $\iota_F$ is also a left inverse and $0$ is also a right unit, so $F$ gives $R[[x]]$ the structure of a group object in the category of topological rings; for two continuous morphisms $f_1,f_2$ from $R[[x]]$ to some ring $A$, given by sending $x$ to $a_1,a_2$ respectively, we define $(f_1 \oplus f_2)(x) = F(a_1,a_2)$, and by our considerations this makes $\Hom_{\text{cont}}{R,A}$ into a group, functorially in $A$.

@@

\subsection{From \texorpdfstring{$C(\Z_p)$}{C(Z\_p)} to \texorpdfstring{$J(\Z_p)$}{J(Z\_p)}}
\label{subsection:czptojzp}
By Subsection \ref{subs:smoothzppoints}, we also know that $C(\Z_p)_P$ is in bijection with $\Z_p$, again with the bijection given by evaluating a parameter and dividing by $p$. The resulting function $\Z_p \to \Z_p^g$ is linear and non-constant modulo $p$, i.e. there are power series $f_1,\dots,f_{g-1} \in \Z_p\angle{z_1,...,z_g}$ such that the image of $C(\Z_p)_P$ is exactly given by $V(f_1,...,f_{g-1})$, and all $f_i$ are linear modulo $p$. Another way to think of this, is as $C(\Z/p^2\Z)_P$ being an affine line inside $J(\Z/p^2\Z)_t$.
TODO: give proof. This uses the functoriality of smooth $\Z_p$-points reducing to a point. That $C(\Z/p^2\Z)_P$ is an affine line inside $J(\Z/p^2\Z)_t$ can also be seen more easily, using Lemma~\ref{lem:smoothpointsfunc}. 
% Say something about the f_i defining C(Z_p) in J(Z_p)

\subsection{Finding \texorpdfstring{$C(\Z)$}{C(Z)}}
Clearly, as subsets of $J(\Z_p)_t$, we have the inclusion $C(\Z)_P \subset C(\Z_p)_P \cap \overline{J(\Z)_t}$. The theory we have built so far enables the following method, which is in sharp contrast with Colemans method; instead of pulling back equations for $\overline{J(\Z)_t}$ to $C(\Z_p)_P$, we pull back equations for $C(\Z_p)_P$ to $\overline{J(\Z)_t}$ to arrive at the following theorem.
\begin{thm}
Let $C,J,p,P,t$ be as in Section~\ref{section:linchab}, let $\kappa$ be as in Theorem~\ref{thm:kappanice} and let $f_1,\cdots,f_{g-1}$ be as in Subsection~\ref{subsection:czptojzp}. Define $\lambda_1 = \kappa^*f_1,\dots,\lambda_{g-1} = \kappa^*f_{g-1}$ to be the pullbacks along $\kappa$ of the $f_i$. Then $\kappa$ induces a surjection from $Z(\lambda_1,\dots,\lambda_{g-1})$ to $C(\Z_p)_P \cap \overline{J(\Z)_t}$.
\end{thm}

The $\lambda_i$ are in general not computationally available, as they are power series, but we can approximate them modulo powers of $p$. For example, we know that the $\lambda_i$ are linear modulo $p$. And in particular, if $f_1,\dots,f_{g-1}$ are as in Remark~\ref{rem:fislinear}, then $\lambda_i = \kappa_i$ for $i \leq g-1$, so they are also of degree at most $m$ modulo $p^m$ for $m < p-1$. Knowing this, we can compute $\lambda_i$ modulo powers of $p$ by extrapolation, and the following lemma, an immediate consequence of Lagrange interpolation, tells us how.
\begin{lem}
\label{lem:interpolate}
Let $R$ be a ring, and assume operations in $R$ are $O(1)$, $f \in R[x]$ with $x = (x_1,\dots,x_n)$ a polynomial of degree at most $m$, and $S$ a subset of $R$ with $S-S \subset \{0\} \cup R^*$ and $|S| > m$. Then $f$ is uniquely determinable in $O(nm^{n+1})$ by its values on $S^n$.
\end{lem} %complexity easily 
\begin{proof}
Both uniqueness and construction follow from the case with $n=1$; uniqueness from the fact that the ideals $\{(x-s) \mid s \in S\}$ are pairwise coprime, and the construction is Lagrange interpolation. The complexity follows easily from induction on $n$ and the formula for Lagrange interpolation.
\end{proof}

@@

\section{Complications and improvements}
\begin{itemize}
\item $\overline{J(\Z)_0}$ open in $J(\Z_p)_0$.
\item Calculations in $J(\Z)$.
\item According to McCallum\&Poonen, it can be that $J(\Z)_0$ and $C(\Z)_P$ are tangent as $p$-adic manifolds and then finite-precision calculations will never be enough. I don't see what's happening there. Maybe give an example? 
\item It is possible that $C(\Z_p)_P \cap \overline{J(\Z)_t}$ is bigger than $C(\Z)_P$; give an example. Refer to \citep{Balakrishnan19}.
\item Explain Mordell-Weil sieve. Maybe "The Mordell-Weil sieve: Proving non-existence of rational points on curves" for a short explanation? Nice to also treat this in the final example.
\item in general, generators of the Mordell-Weil group are computationally out of reach. But, if we only generate a subgroup of $J(\Z)_0$ of index finite and coprime to $p$, then the closure of this in $J(\Z_p)_0$ will be the same. This happens in particular if we generate a subgroup of $J(\Z)$ of index finite and coprime to $p|J(\F_p)|$. Preferably, find a source for this?
\end{itemize}

\subsection{Calculations modulo \texorpdfstring{$p^2$}{p\^2}}
Say something about how, for the right set of points $\{p_1,...,p_g\}$, the map $C^{g} \to J$ is etale at that set of points. Then say something about what that implies about $J(\Z_p)_0$ (equal to $\sum q_i - p_i$ where $q_i \in C(\Z_p)_{p_i}$), and $J(\Z/p^2\Z)_0$, which becomes isomorphic as vector spaces to $\F_p^g$, and how we can find $\kappa$ by expressing elements of $J(\Z)_0$ in that basis, and similarly $f_1,...,f_{g-1}$ by finding $P_1 - P_2$ in that basis for some $Q_1,Q_2 \in C(\Z/p^2\Z)_P$ not equal.

An important part of this, is that there is a parametrisation $P_\mu, \mu \in \F_p$ of the points in $C(\Z/p^2\Z)_P$ such that $P_\mu + P_\nu = P_{\mu + \nu} + P_0$ as Cartier divisors.

\section{Implementations of linear Chabauty and an explicit example}
[TODO: make sure I clearly differentiate divisors and divisor classes]
We now assume that $C$ is hyperelliptic, i.e. given by the degree $2g+2$ homogenisation of an equation of the form
\[
y^2 = f(x)
\]
inside $\P(1,g+1,1)$ where $f$ is a monic polynomial of degree $2g+1$ or $2g+2$. An alternative way of defining such a curve, and the one we will be using mainly, is as a glueing of two affine charts: $y^2 = f(x)$, and $w^2 = f^{r}(v)$, where $f^{r}(v)$ is the polynomial $v^{2g + 2} f(1/v)$, and a birational map between them is given by $(x,y) \mapsto (\frac{1}{x},\frac{y}{x^{-g-1}})$. We also have the coordinates $X,Y,Z$ of $\P(1,g+1,1)$, with $x = X/Z, y = Y/Z^{g+1}, v = Z/X, w = Y/X^{g+1}$, but beware; these coordinates do not behave nicely on the origin of the patch $D(Y)$. We mainly use the first chart; we call any point that lies on it an affine point of $C$. We will treat the case that $f$ has degree $2g+2$ (this can be done by translating $f$ until the constant coefficient is non-zero, and then looking at $f^r$). In that case, near the line at infinity $C$ looks like $Y^2 = X^{2g+2}$, i.e. $(Y-X^{g+1})(Y+X^{g+1}) = 0$, and we see there are two points $\infty_+ = (1:1:0)$ and $\infty_- = (1:-1:0)$. Finally, we note that there is an involution on $C$ given by $\sigma(x,y) = (x,-y)$ and $\sigma(v,w) = (v,-w)$.

@@

We first calculate $\Lcal(2D_0 - D_{\text{aff}})$. Note that \todo{Compute this, currently talking with Mascot about this.}

@@

Say something about how one express an element in $J(\Z/p^2\Z)_0$ on the basis in something like $O(g)$ hopefully (ignoring the time for adding,multiplying, et cetera in the Jacobian), identify $J(Z/p^2)_0$ with $F_p^g$ using $g$ generators of the ideal of the zero section of $J_{Z/p^2}$, then the group structure that $J$ induces on $F_p^g$ is the usual addition. This is currently done using code from Mascot. Say what the final complexity is.

>>>>

\usepackage{graphicx}
%\pagestyle{empty}
\usepackage{dsfont}
\usepackage{textcomp}
\usepackage{amsmath}
\usepackage{courier}
\usepackage[utf8]{inputenc}
\usepackage{graphicx}
\graphicspath{{images/}}
\usepackage{amsthm}
\usepackage{eucal}
\usepackage{amssymb}
%\usepackage{mdsymbol}
\usepackage{mathrsfs}
\usepackage{dsfont}
\usepackage[usenames,dvipsnames,svgnames,table]{xcolor}
\usepackage{caption}
\usepackage{tikz}
\usepackage{tikz-cd}
\usepackage{wrapfig,lipsum,booktabs}
\usepackage{caption}
\usepackage[square, longnamesfirst,numbers]{natbib}
\usepackage[ruled,linesnumbered]{algorithm2e}
\usepackage{todonotes}
\usepackage[all]{nowidow}

@@

{\LARGE\bf 
A geometric approach to linear Chabauty
}

@@

This Coleman-Chabauty method has been greatly generalised by Kim to so called non-abelian Chabauty. He interprets working in the Jacobian as dealing with the abelianised fundamental group of the curve $C$, and works with larger, non-abelian quotients of the fundamental group. Using quadratic Chabauty, a special case of non-abelian Chabauty, Balakrishnan, Dogra, M\"uller, Tuitman and Vonk famously mcalculated in May of 2019 all rational points of the ``cursed curve'', the modular curve $X_s(13)$. They use an endomorphism of the Jacobian, and do $p$-adic analysis on $p$-adic local heights to make non-abelian Chabauty explicit. \todo{Bas, is dit de goede samenvatting? Voornamelijk mijn vraag: is het nog steeds een specifiek geval van Coleman-Chabauty?}

Very recently, Edixhoven and Lido made available a preprint of their article ``Geometric quadratic Chabauty''. Their goal in this article is to make a more geometric approach to effective quadratic Chabauty, by working in a pullback $T$ of the Poincar\'e torsor of the Jacobian. They also abandon the Coleman method for Chabauty; instead they parametrise the map $T(\Z) \to T(\Z_p)$ with power series, and pull back equations for the curve along this map.

@@

\begin{defn}
Let $S$ be a scheme, $T \to U$ a morphism of schemes and $x : T \to S$ a $T$-point. We define $S(U)_x$ as the morphisms from $U$ to $S$ that, after precomposing with $T \to U$, give $x$.
\end{defn}
\begin{exmp}
If we have a projective variety $X$ over $\Z$, then $X(\Z)$ is in bijection with $X(\Q)$. The natural map $X(\Z) \to X(\F_p)$ reduces a point modulo $p$ and for $x \in X(\F_p)$, the set $X(\Z)_x$ consists of the $\Z$-points reducing to $x$.
\end{exmp}

Geometric chabauty works by finding an upper bound for the cardinality of $C(\Z)_P$. For this, we use the following diagram.
\[
\begin{tikzcd}
C(\Z)_P \arrow[d] \arrow[r, "a"] & J(\Z)_t \arrow[d] \\
C(\Z_p)_P \arrow[r]              & J(\Z_p)_t        
\end{tikzcd}
\]
where the two vertical maps are inclusions, and the two horizontal maps are subtraction of $b$. We will in fact compute upper bounds for the larger set $C(\Z_p)_P \cup \overline{J(\Z)_t}$, where $\overline{J(\Z)_t}$ is the closure of $J(\Z)_t$ in $J(\Z_p)_t$.

In Section~\ref{section:smoothzppoints}, we treat the structure of $C(\Z_p)_P$ and $J(\Z_p)_t$; we will show that, after choosing parameters, they are canonically in bijection with respectively $\Z_p$ and $\Z_p^g$. We also discuss the resulting map $\Z_p \to \Z_p^g$.

In Section~\ref{section:kappa} we further look at the map $J(\Z)_t \to J(\Z_p)_t$. This is a translation of the group morphism $J(\Z)_0 \to J(\Z_p)_0$ between kernels of reduction; it turns out that for our choice of $p$, the subgroup $J(\Z)_0$ is free of rank $r$. We study the properties of the resulting map $\Z^r \to \Z_p^g$, using the theory of formal groups to determine the group structure on $\Z_p^g$ induced by the bijection $J(\Z_p)_t \to \Z_p^g$.

In Section~\ref{section:intersection}, we put all of this information together, culminating in several methods to compute upper bounds on $\left|C(\Z_p)_P \cup \overline{J(\Z)_t}\right|$ with finite precision calculations.

These methods are not always guaranteed to prove finiteness of $C(\Z)_P$. In Section~\ref{section:remarks} we treat several possible complications and improvements, partially with explicit examples. We also look at the specific case of calculations modulo $p$, i.e. in $J(\Z/p^2\Z)_t$, in which the theory simplifies. \todo{Bas, wat vind je van het bestaan van dit hoofdstuk?}

Next, in Section~\ref{section:explicit} we focus on how to perform these calculations happening in the Jacobian. Here, we use Makdisi's approach of representing a divisor by a subspace of a large Riemann-Roch space. We partially follow the article \citep{mascot2018} by Mascot, and also give an original algorithm for explicitly computing the map $C(\Z_p)_P \to J(\Z_p)_t$ in Makdisi's representation.

We end with an explicit example in Section~\ref{section:example}, a genus $2$ curve whose Jacobian has Mordell-Weil rank $1$. We use both Magma and Pari/GP to do our calculations, and end up with a complete list of all rational points of the curve.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\newpage
% Say something about smooth schemes over Z_p, and their Z_p points
\section{Points on a smooth scheme over \texorpdfstring{$\Z_p$}{Z\_p}}
\label{section:smoothzppoints}
Let $X/\Z_p$ be a smooth scheme of relative dimension $d$, and let $x \in X(\F_p)$ be a point. Then $X(\Z_p)_x$ is in bijection with $\Z_p^d$. This bijection is given by choosing parameters at $x$; evaluating at $X(\Z_p)_x$ gives a bijection with $(p\Z_p)^d$, and then we divide by $p$. For putting up a nice framework to work in, we start with blowing up $X$ at $x$.
%TODO: give proof and define $\Z_p\angle{z_1,...,z_r}$ as the $p$-adic completion of $\Z_p[z_1,...,z_r]$, i.e. powerseries which are polynomials modulo every power of $p$.

@@

\subsection{From \texorpdfstring{$C(\Z_p)$}{C(Z\_p)} to \texorpdfstring{$J(\Z_p)$}{J(Z\_p)}}
\label{subsection:czptojzp}
We know that $C(\Z_p)_P$ is in bijection with $\Z_p$, again with the bijection given by evaluating a parameter and dividing by $p$. The resulting function $\Z_p \to \Z_p^g$ is linear and non-constant modulo $p$, i.e. there are power series $f_1,\dots,f_{g-1} \in \Z_p\angle{z_1,...,z_g}$ such that the image of $C(\Z_p)_P$ is exactly given by $V(f_1,...,f_{g-1})$, and all $f_i$ are linear modulo $p$. Another way to think of this, is as $C(\Z/p^2\Z)_P$ being an affine line inside $J(\Z/p^2\Z)_t$.
TODO: give proof. This uses the functoriality of smooth $\Z_p$-points reducing to a point. That $C(\Z/p^2\Z)_P$ is an affine line inside $J(\Z/p^2\Z)_t$ can also be seen more easily, using Lemma~\ref{lem:smoothpointsfunc}. 
% Say something about the f_i defining C(Z_p) in J(Z_p)

\section{From \texorpdfstring{$J(\Z)$}{J(Z)} to \texorpdfstring{$J(\Z_p)$}{J(Z\_p)}}
\label{section:kappa}
For $p > 2$, we know that the torsion of $J(\Z)$ injects into $J(\F_p)$, by Proposition 2.3 of \cite{pierre2000}. Hence for $0 \in J(\F_p)$, we know $J(\Z)_0$ is as a group isomorphic to $\Z^r$ with $r$ the Mordell-Weil rank. By assumption, we also know $J(\Z)_t$ is in bijection with $J(\Z)_0$, with the bijection giving by translating with a lift of $t$. By Subsection \ref{section:smoothzppoints} we know $J(\Z_p)_t$ is in bijection with $\Z_p^{g}$, with the bijection given by evaluating parameters and dividing by $p$. Let $\kappa: \Z^r \to \Z_p^g$ be the map resulting from the inclusion $J(\Z)_t \to J(\Z_p)_t$. Then $\kappa$ turns out to have a special property.

@@

\subsection{From group schemes to formal groups}
We first recall the definition of a formal group. For this entire section, $R$ is a ring, and $x,y,z$ are vectors of variables.
\begin{defn}
\label{defn:formalgroup}
Let $n$ be a non-negative integer. Let $x,y,z$ be vectors of $n$ variables. An $n$-dimensional formal group is an element $F = (F_1,\dots,F_n) \in R[[x,y]]_0^n$ satisfying $F \equiv x + y \mod (x,y)^2$ and $F(F(x,y),z) = F(x,F(y,z))$. If furthermore $F(x,y) = F(y,x)$, this formal group is said to be commutative.
\end{defn}
\begin{exmp}
\label{exmp:gmformalgroup}
Take $n=1$ and $F(x,y) = x + y + xy = (1+x)(1+y)-1$, also known as the multiplicative formal group. This satisfies associativity as $F(F(x,y),z) = (1+x)(1+y)(1+z)-1 = F(x,F(y,z))$.
\end{exmp}
\begin{exmp}
\label{exmp:ganformalgroup}
For any $n$, we can take $F(x,y) = x+ y$, also known as the $n$-dimensional additive formal group.
\end{exmp}
[TODO: add text explaining the following lemma]
Note there is no mention of an inverse, but the following lemma, a formal version of the implicit function theorem, tells us that the inverse follows automatically from the definitions.
\begin{lem}
\label{lem:implicitfunction}
Let $x,y$ have length $n$. Let $F \in R[[x,y]]_0^n$ such that $F \equiv Ax + By \bmod (x,y)^2$ with $B \in \GL_n(R)$. Then there's a unique $\iota \in R[[x]]_0^n$ such that $F(x,\iota(x)) = 0$. 
\end{lem}
\begin{proof}
If $F$ is a polynomial, this is a special case of a multivariate version of Hensel's lemma, as in Corrolaire~2 of \cite[III,4.5]{BourbakiCA}, over $R[[x]]$ with maximal ideal $\m = (x)$, as the derivative matrix of $F(x,\iota)$ with respect to $\iota$ is $B$, which is invertible, and $\iota = 0$ gives a solution modulo $\m$. Now for $F$ any power series, let $F_j$ be the polynomial consisting of all terms in $F$ of degree at most $j$, and let $\iota_j$ be the unique power series in $R[[x]]_0^n$such that $F_j(x,\iota_j) = 0$. Note that both $\iota_j$ and $\iota_{j+1}$ are solutions to $F_j(x,\iota) \equiv 0 \mod \m^j$, so by the uniquess guaranteed by Hensel's lemma used over $R[[x]]/\m^j$, these must be equal. Hence they converge to $\iota \in R[[x]]_0^n$, which is the unique solution of $F(x,\iota(x)) = 0$.
\end{proof}
This has the following corollary about the inverse of a power series.
\begin{cor}
\label{lem:formalinverse}
Let $x$ have length $n$. Let $a \in R[[x]]_0^n$ with $a \equiv Px \mod x$ for some matrix $A\in \GL_n(R)$. Then there is a unique $b \in R[[x]]_0^n$ such that $a \circ b = b \circ a = x$.
\end{cor}
\begin{proof}
Let $F(x,y) = x - a(x)$. This satisfies the constraints of Lemma~\ref{lem:implicitfunction}, so we find a unique $b$ such that $a \circ b = x$. Applying Lemma~\ref{lem:implicitfunction} again gives a unique $c$ such that $b \circ c = x$. But then $a = a \circ (b \circ c) = (a \circ b) \circ c = c$ shows $a = c$ and hence we are done.
\end{proof}
So a formal group $F$ does by Lemma~\ref{lem:implicitfunction} indeed have a right inverse $\iota_F$. Also, by $F(0,0) = 0$ we have that $F(x,F(\iota_f,0)) = 0$ so $F(\iota_f,0)$ is in fact equal to $\iota_F$; as $\iota_F \equiv -x \mod (x)^2$, it has a formal inverse and hence $F(x,0) = x$, so $0$ is indeed a right unit. A standard argument now shows that $\iota_F$ is also a left inverse and $0$ is also a right unit, so $F$ gives $R[[x]]$ the structure of a group object in the category of topological rings; for two continuous morphisms $f_1,f_2$ from $R[[x]]$ to some ring $A$, given by sending $x$ to $a_1,a_2$ respectively, we define $(f_1 \oplus f_2)(x) = F(a_1,a_2)$, and by our considerations this makes $\Hom_{\text{cont}}{R,A}$ into a group, functorially in $A$.

@@

\section{Computing the intersection}
\label{section:intersection}
Clearly, as subsets of $J(\Z_p)_t$, we have the inclusion $C(\Z)_P \subset C(\Z_p)_P \cap \overline{J(\Z)_t}$. The theory we have built so far enables the following method, which is in sharp contrast with Colemans method; instead of pulling back equations for $\overline{J(\Z)_t}$ to $C(\Z_p)_P$, we pull back equations for $C(\Z_p)_P$ to $\overline{J(\Z)_t}$ to arrive at the following theorem.
\begin{thm}
Let $C,J,p,P,t$ be as in the overview, let $\kappa$ be as in Theorem~\ref{thm:kappanice} and let $f_1,\cdots,f_{g-1}$ be as in Subsection~\ref{subsection:czptojzp}. Define $\lambda_1 = \kappa^*f_1,\dots,\lambda_{g-1} = \kappa^*f_{g-1}$ to be the pullbacks along $\kappa$ of the $f_i$. Then $\kappa$ induces a surjection from $Z(\lambda_1,\dots,\lambda_{g-1})$ to $C(\Z_p)_P \cap \overline{J(\Z)_t}$.
\end{thm}

The $\lambda_i$ are in general not computationally available, as they are power series, but we can approximate them modulo powers of $p$. For example, we know that the $\lambda_i$ are linear modulo $p$. And in particular, if $f_1,\dots,f_{g-1}$ are as in Remark~\ref{rem:fislinear}, then $\lambda_i = \kappa_i$ for $i \leq g-1$, so they are also of degree at most $m$ modulo $p^m$ for $m < p-1$. Knowing this, we can compute $\lambda_i$ modulo powers of $p$ by extrapolation, and the following lemma, an immediate consequence of Lagrange interpolation, tells us how.
\begin{lem}
\label{lem:interpolate}
Let $R$ be a ring, and assume operations in $R$ are $O(1)$, $f \in R[x]$ with $x = (x_1,\dots,x_n)$ a polynomial of degree at most $m$, and $S$ a subset of $R$ with $S-S \subset \{0\} \cup R^*$ and $|S| > m$. Then $f$ is uniquely determinable in $O(nm^{n+1})$ by its values on $S^n$.
\end{lem}
\begin{proof}
Both uniqueness and construction follow from the case with $n=1$; uniqueness from the fact that the ideals $\{(x-s) \mid s \in S\}$ are pairwise coprime, and the construction is Lagrange interpolation. The complexity follows easily from induction on $n$ and the formula for Lagrange interpolation.
\end{proof}

@@

\section{Complications and improvements}
\label{section:remarks}
\begin{itemize}
\item $\overline{J(\Z)_0}$ open in $J(\Z_p)_0$.
\item Calculations in $J(\Z)$.
\item According to McCallum\&Poonen, it can be that $J(\Z)_0$ and $C(\Z)_P$ are tangent as $p$-adic manifolds and then finite-precision calculations will never be enough. I don't see what's happening there. Maybe give an example? 
\item It is possible that $C(\Z_p)_P \cap \overline{J(\Z)_t}$ is bigger than $C(\Z)_P$; give an example. Refer to \citep{Balakrishnan19}.
\item Explain Mordell-Weil sieve. Maybe "The Mordell-Weil sieve: Proving non-existence of rational points on curves" for a short explanation? Nice to also treat this in the final example.
\item in general, generators of the Mordell-Weil group are computationally out of reach. But, if we only generate a subgroup of $J(\Z)_0$ of index finite and coprime to $p$, then the closure of this in $J(\Z_p)_0$ will be the same. This happens in particular if we generate a subgroup of $J(\Z)$ of index finite and coprime to $p|J(\F_p)|$. Preferably, find a source for this? Bas, do you have a reference for the complexity of such calculations?
\end{itemize}

\section{Implementations of linear Chabauty}
We now assume that $C$ is hyperelliptic, i.e. given by the degree $2g+2$ homogenisation of an equation of the form
\[
y^2 = f(x)
\]
inside $\P(1,g+1,1)$ where $f$ is a monic polynomial of degree $2g+1$ or $2g+2$. An alternative way of defining such a curve, and the one we will be using mainly, is as a glueing of two affine charts: $y^2 = f(x)$, and $w^2 = f^{r}(v)$, where $f^{r}(v)$ is the polynomial $v^{2g + 2} f(1/v)$, and a birational map between them is given by $(x,y) \mapsto (\frac{1}{x},\frac{y}{x^{-g-1}})$. We also have the coordinates $X,Y,Z$ of $\P(1,g+1,1)$, with $x = X/Z, y = Y/Z^{g+1}, v = Z/X, w = Y/X^{g+1}$, but beware; these coordinates do not behave nicely on the origin of the patch $D(Y)$. We mainly use the first chart; we call any point that lies on it an affine point of $C$. We will treat the case that $f$ has degree $2g+2$ (this can be done by translating $f$ until the constant coefficient is non-zero, and then looking at $f^r$). In that case, near the line at infinity $C$ looks like $Y^2 = X^{2g+2}$, i.e. $(Y-X^{g+1})(Y+X^{g+1}) = 0$, and we see there are two points $\infty_+ = (1:1:0)$ and $\infty_- = (1:-1:0)$. Finally, we note that there is an involution on $C$ given by $\sigma(x,y) = (x,-y)$ and $\sigma(v,w) = (v,-w)$.

@@

We first calculate $\Lcal(2D_0 - D_{\text{aff}})$. Both over $\Q_p$ and over $\F_p$ we have the equality
\[
\Lcal(2D_0 - D_{\text{aff}) = a(x)\Lcal((2g+2-d)(\infty_- + \infty_+)) + (y-b(x))\Lcal((g+1)(\infty_- + \infty_+))
\]
and then it should follow over $\Z_p$ so over $\Z/p^k\Z$. \todo{Give more details, currently talking with Mascot about this.}

@@

Say something about how one express an element in $J(\Z/p^2\Z)_0$ on the basis in something like $O(g)$ hopefully (ignoring the time for adding,multiplying, et cetera in the Jacobian), identify $J(\Z/p^2)_0$ with $\F_p^g$ using $g$ generators of the ideal of the zero section of $J_{\Z/p^2}$, then the group structure that $J$ induces on $\F_p^g$ is the usual addition. This is currently done using code from Mascot.

\subsection{Complexity}
Say what the final complexity is.
